A Java Wrapper For Embarrassingly Parallel Problems

### 2013-11-01 paper/paper.tex
More paper work

\begin{abstract}
An embarrassingly parallel problem, also called pleasingly parallel, is one that 
can easily be broken up into identital components that do not need to interact 
to find the solution. Several problems in the biological sciences are 
embarrsingly parallel but they are difficult to rewrite to use multithreading, 
because they are either poorly written or too complex. We have written a 
wrapper, implemented in Java, that will execute the serial program in several 
threads. RESULTS SENTENCE. More
statistics, savable configurations and more options will be implemented next, in
addition to better testing and refactored.
\end{abstract}

@@

The Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. The Basic Local
Alignment Search Tool (BLAST) \cite{BLAST} is a common program used to find
similarities between sequences and then compares them to databases to calculate
the statistical significance. Another program we have been looking at is called
Trinity. Trinity claims to be an efficient way to do de novo reconstruction of
transcriptomes from mRNA sequence data. With this variety of applications that 
would benefit from parallelism, it was obvious it would be desirable to develop
a framework for creating wrapper programs.

@@

In this version of the wrapper, several components are being added. First, there
is a wizard style menu that will allow the end-user to select not only the 
program that wish to use, but also to manage other settings for the run. A new 
IO class splits the input and merges the output more efficiently. Lastly, in 
order to look at how well the wrapper performs, the threads keep their own 
statistics, which will be used to evaluate the effectiveness of the framework 
and to find potential enhancements and help end-users determine whether their 
use of the framework is efficient.

@@

Tables charts, etc

@@

Conclusions gathered from results.

>>>>

\begin{abstract}
An embarrassingly parallel problem, also called pleasingly parallel, is one that 
can easily be broken up into identital components that do not need to interact 
to find the solution. Several problems in the biological sciences are 
embarrsingly parallel but they are difficult to rewrite to use multithreading, 
because they are either poorly written or too complex. We have written a 
framework, implemented in Java, that will execute the serial program in several 
threads. These threads working together help cut down the total runtime
exponentially. More statistics, savable configurations and more options will be 
implemented next, in addition to better testing and refactored.
\end{abstract}

@@

REDO THIS PART
The Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. The Basic Local
Alignment Search Tool (BLAST) \cite{BLAST} is a common program used to find
similarities between sequences and then compares them to databases to calculate
the statistical significance. Another program we have been looking at is called
Trinity. Trinity claims to be an efficient way to do de novo reconstruction of
transcriptomes from mRNA sequence data. With this variety of applications that 
would benefit from parallelism, it was obvious it would be desirable to develop
a framework for creating wrapper programs.

@@

In this version of the framework, several components are being added. First, 
there is a wizard style menu that will allow the end-user to select not only the 
program that wish to use, but also to manage other settings for the run. A new 
IO class splits the input and merges the output more efficiently. Lastly, in 
order to look at how well the framework performs, the threads keep their own 
statistics, which will be used to evaluate the effectiveness of the framework 
and to find potential enhancements and help end-users determine whether their 
use of the framework is efficient.

@@

Table and chart

The Ka/Ks Calculator was run through the framework with ten different samples,
each containing sixteen units of work. Each sample was run by the framework at 
five different level of numbers of worker threads, starting at 1 and 
increasing by powers of two until the 100\% CPU usage (16 processors). As the 
numbers of threads increased, the time it took descreased exponentially.

@@

In a few cases, there was an increase in time when attempting to max out the CPU 
usage. This occured because the framework had to wait until for the CPU to be 
finished with its previous work before it was allowed to start work on what the 
framework wanted it to do. In the cases when the framework executed faster while
the CPU was maxed out, it was only faster a minute at the most. With such a
small time gain with the CPU at 50\% versus the CPU at 100\%, running the
framework at 50\% CPU is more efficient when the number of threads is on the
same order as the number of units of work.
